% Acknowledgments
The current text is the result of my PhD studies. The academic process has
taken place in Dionysos Satellite Observatory, at the School of Rural, Surveying 
and Geoinformatics Engineering, of the National Technical University of Athens 
(NTUA).

Firstly, I would like to express my sincere gratitude to my supervisor, Prof. Maria 
Tsakiri. Her guidance, support and assistance, not only on technical matters, has 
been of major importance during the course of my PhD. Without her endorsement, i 
would have never concluded this task. Furthermore, my gratitude goes to Emeritus Prof. 
Demitris Paradissis. Since my graduation, Prof. Paradissis has been a scientific 
mentor to me and has lavishly shared his knowledge and understanding throughout 
the nearly 15 years of out collaboration. 

I would like to thank all of my friends and colleagues at Dionysos Satellite 
Observatory; Jordan Galanis, Thanasis Zisopoulos and Kostas Raptakis. Way to 
often have they transformed dull working hours in times to be remembered. Special 
thanks goes to Vangelis Zacharis, who apart from supporting me and sharing his 
expertise, has always been ungrudgingly listening to my mumbling about bugs and 
km-level errors that shouldn't be there. The same goes for Dr. Pol Kolokousis, who 
pretty much taught me how computers work.

It has been with great pleasure that throughout this journey I have been standing 
along my good friend Prof. Demitris Anastasiou. His hard work, commitment and 
progress have been a joy to watch and an example to strive for. His help and 
support were simply overwhelming.

%A big thank-you is in due to Dr. Samuel Nahmani and Dr. Arnaud Pollet. They had no 
%second thoughts in supporting what seemed like an overreaching, pretty ``crazy'' idea 
%a few years back. Since then, they have provided crucial feedback and support in 
%a number of ways.
A big thank-you is in due to Dr. Samuel Nahmani and Dr. Arnaud Pollet. They showed no 
glimpse of fear (!) when a few years back I showed up to them proposing an overreaching,
pretty ``crazy'' (to be honest) idea. Since then, they have provided crucial 
feedback and support.

The warm welcome I received from the \gls{ids} has also played a big role in the 
fulfillment of this Thesis. I would especially like to thank Guilhem Moreaux, 
head of the \gls{ids} Combination Center, Petr Štěpánek, \gls{ids} Analysis 
coordinator, Jérôme Saunier of IGN and Jean-Michel Lemoine of CNES. 
Dr. Martin Lasser of the Astronomical 
Institute, University of Bern, has also been very kind in providing all information 
i needed and answering all questions i posed regarding validation of results.

I would like to thank my friends and family, and especially my parents
Lena and Thanasis for their unconditional, ever lasting help and support in more
ways i could ever count.

Last but not least, my gratitude goes above all to \textgreek{\emph{Αλεξάνδρα}}
who had to put up with my absent-mindedness for quite some time! Thank you for not 
divorcing me, i will now go put these shelves in place.

Since i am not a big fan of proprietorship, i consider the result of this work to 
be collective yet property of none.
\par
\begin{flushright}
  Xanthos, \today\par
\end{flushright}
\vfill
