\section{The COST-G Benchmark Test}\label{sec:costg-benchmark-test}

The International Combination Service for Time-variable Gravity Fields (COST-G)\footnote{https://cost-g.org/} 
is a product center of the International Gravity Field Service (IGFS) and is dedicated 
to the combination of monthly global gravity field models. In the framework of 
COST-G, gravity field solutions from different analysis centres are combined to 
provide a consolidated solution of improved quality and robustness to users. To 
achieve its goal, the individual products must be of utmost accuracy, using 
state-of-the-art models. 

In order to assess the consistency level of the implemented, underlying models 
from the contributing analysis centers, a benchmark test was devised. 
A comprehensive dataset containing commonly used forces in orbit and gravity 
field modeling, such as Earth's gravity field and tides, was compiled and 
evaluated over a one-day orbit arc of \gls{grace}. Supplementary data were also 
included to facilitate straightforward comparisons. This benchmark test is 
intended to be used as a reference data set and provide the opportunity to test the 
implementation of these models at various institutions involved in orbit and 
gravity field determination from satellite tracking data. The benchmark test is 
described in \cite{Lasser2020}.

This benchmark test is used throughout this document, to evaluate the implementation 
of various models and forces used both here and in the analysis performed to 
derive COST-G products by a series of high quality scientific software packages. Validation of results aimed to comply with the highest standards is hard to achieve, and such 
tests can provide crucial feedback and help identify and mitigate error and bug that would otherwise pass unnoticed.
