\tikzstyle{every picture}+=[remember picture]
\everymath{\displaystyle}

%\section{Questions}
%\begin{enumerate}
%  \item Orbit integration is performed without using (at all) the orbital 
%    elements (numerical integration is performed in GCRS Cartesian coordinates). 
%    Is this approach correct? (we only use state to orbital elements 
%    transformation to check/validate results).
%
%  \item Orbit integration to time $t_i$ results in the extrapolated state 
%    vector and the corresponding state transition matrix $\bm{\Phi}$. The latter 
%    is used to update the filter's var-covar matrix via $\bm{\Phi} \cdot \bm{P} \cdot \bm{\Phi}^T$,
%    but since the integration is performed in GCRS, should we transform the 
%    $\bm{\Phi}$ matrix wrt ECEF?
%
%  \item I use TAI (transformed from proper time using the correction recorded in 
%    the RINEX file) to perform orbit integration; is this correct ? Should i
%    use TCG ? (see \ref{ssec:coordinate-proper-time})
%
%  \item According to \cite{lemoine-2016} (Sec. 2.5.1) a correction of \emph{aberration} 
%    has to ber performed in the observation equation. This is currently not 
%    implemented. How large is its effect? How exactly should it be applied?
%
%  \item Is the formula for the ionospheric correction \ref{eq:dion-diff} (for 
%    the range-rate) correct ?
%
%  \item Ionospheric correction via \ref{eq:iono-delay-cycles} does not take 
%    into account the RINEX record ``\texttt{L2/L1 DATE OFFSET}''. Should it 
%    be taken into account ?
%
%  \item Is the relativistic correction \ref{eq:drel-diff} (for range-rate) 
%    correct ?
%
%  \item Solar radiation pressure is currently ignored. What is the practice 
%    for implementing shadow models (for LEOs) ?
%
%  \item For Jason-3 attitude, we use \emph{body} but not \emph{solar-panel} 
%    quaternions. How do i use the \emph{solar-panel} quaternions ? (see \ref{sssec:ja3-quaternions})
%
%  \item I use the IDS distributed log files to get beacon eccentricities. Is 
%    this good practice ? Should i use the SINEX files ? (see \ref{ssec:beacon_coordinates})
%
%  \item Starting from \ref{eq:lem13}, we need the partials wrt the estimated 
%    parameters. Are \ref{eq:partials-pos} and \ref{eq:partials-vel} correct ?
%
%  \item Is it good practice to use obsevration flags (from RINEX) to mark/exclude 
%    observations ? (see \ref{sssec:rnx-flags})
%
%\end{enumerate}


\section{Observation Model Description}
\label{sec:doris-introduction}

For the DORIS observatrion equation model, we use the formula described in 
\cite{lemoine-2016}, Eq. 17:
\begin{subequations} \label{eq:lem13}
    \begin{align}
        v_{measured} & = \frac{c}{f_{e_N}} (f_{e_N} - f_{r_T} -
          \frac{N_{DOP}}{\Delta\tau_r}) + \Delta u_{REL} + 
          \Delta v_{IONO} \label{eq:lem13a} \\
        v_{theo} &= \frac{\rho_2 - \rho_1}{\Delta\tau_r} +
%         (1- \frac{U_e}{c^2} - \frac{{V_e}^2}{2 c^2}) + 
          \Delta v_{TROPO} - \frac{c(\frac{N_{DOP}}{\Delta\tau_r} + 
          f_{r_T})}{f_{e_N}} \frac{\Delta f_e}{f_{e_N}} \label{eq:lem13b}
    \end{align}
\end{subequations}

where $v_{measured}$ and $v_{theo}$ represent the relative velocity between  
the emitter and the receiver as measured and computed respectively.

\subsection{Nominal Receiver \& Emitter Frequencies}
\label{ssec:nominal-frequencies}
In the observation equation model \ref{eq:lem13} we distinguish between 
\emph{nominal} and \emph{true} receiver/emitter frequencies, to account for 
the fact that in ``real world'' these two are not actually equal.

\subsubsection{Emitter/Beacon Nominal Frequencies, $f_{e_N}$}
\label{sssec:beacon-nominal-frequencies}

RINEX file headers, contain values of the \emph{station frequency shift 
factor} $k$ for each of the beacons involved (\cite{DORISRNX3}, Sec. 6.16). 
These are used to compute the ``nominal'' frequencies of the beacon/emitter. 
(usually, this shift factor is just $0$, but it can be an integer 
$k \neq 0$). The frequencies are computed as (\cite{DORISRNX3}, Sec. 6.16):
\begin{equation}
  \begin{aligned}
    L_{2GHz}   &= 543 \cdot F_0 \left( \frac{3}{4} + \frac{87\cdot k}{5 \cdot 2^{26}} \right) \\
    L_{400MHz} &= 107 \cdot F_0 \left( \frac{3}{4} + \frac{87\cdot k}{5 \cdot 2^{26}} \right) 
    \label{eq:nominal-freq}
  \end{aligned}
\end{equation}
where $F_0 = 5e6 \text{ Hz}$ the USO frequency. These value, are the ones 
labelled as $f_{e_N}$ in \ref{eq:lem13}.

The \emph{true proper frequency} of the emitter $f_{e_T}$, can be computed 
from:
\begin{equation}
  f_{e_T} = f_{e_N} \cdot \left( 1 + \frac{\Delta f_e}{f_{e_N}} \right)
\end{equation}
but the quantity $\Delta f_e / f_{e_N}$ is not know a-priori and has to be 
estimated during the processing.

We estimate $\Delta f_e / f_{e_N}$ using a \colorbox{blue!20}{linear model using proper time}, 
aka $\frac{\Delta f_e}{f_{e_N}}\at{\tau=\tau _i} = \alpha + \beta \cdot \delta \tau$
with an a-priori value of 0 and no process noise. For the estimation, we need 
the partials of the observation equation \ref{eq:lem13b} w.r.t. the $\alpha$ and 
$\beta$ parameters, which are:
\begin{equation}
  \begin{aligned}
  \frac{\partial v_{theo}}{\partial \alpha} &= 
    \frac{c(\frac{N_{DOP}}{\Delta\tau_r} + f_{r_T})}{f_{e_N}} \\
  \frac{\partial v_{theo}}{\partial \beta} &= 
    \frac{c(\frac{N_{DOP}}{\Delta\tau_r} + f_{r_T})}{f_{e_N}} \cdot \delta \tau
  \end{aligned}
\end{equation}

\subsection{Receiver True Proper Frequency $f_{r_T}$}
\label{ssec:receiver-true-proper-frequency}
In equation \ref{eq:lem13}, $f_{r_T}$ is the \emph{true proper frequency of 
the receiver} with nominal value 
\begin{equation}
  f_{r_T} = f_{r_N} \cdot \left( 1 + \frac{\Delta f_r}{f_{r_N}} \right)
\end{equation}
where $f_{r_N}$ is the ``nominal'' frequency value.

In practice the value $\Delta f_r / f_{r_N}$, called the \emph{relative 
frequency offset} of the receiver, is given in the RINEX files for each epoch 
(under the observable tagged \texttt{F}). Note that these values are scaled to 
$10^{-11}$ (\cite{DORISRNX3}, Sec. 6.11), so that for a given epoch $t_i$, the 
true frequency is
\begin{equation}
  f_{r_T}\at{t=t_i} = f_{r_N} \cdot \left( 1 + F_{t_i} \cdot 10^{-11} \right)
  \label{eq:tpf-rec}
\end{equation}
where $F_{t_i}$ is the relative frequency offset value recovered from the RINEX 
file.

In our implementation, we use ``smoothed'' values of RINEX-provided 
$\Delta f_r / f_{r_N}$ estimates. In a first step/pass we use a linear model 
to filter all the values in the RINEX file (as suggested by \cite{lemoine-2016}, 
Sec. 2.5.4). We then use this linear model to compute relative frequency offset 
values when needed. 

\subsection{Beacon Coordinates}
\label{ssec:beacon_coordinates}
Coordinates for the DORIS Network sites, are extracted from the \texttt{dpod*} 
SINEX files (\cite{Moreaux2020}). To get the coordinates of the sites at the 
beacon reference point, we procced as follows:
\begin{enumerate}
  \item extrapolate coordinates (in the DPOD, ECEF reference frame) at the 
    requested epoch, $\bm{r}\sp{\prime}$
  \item get site eccentricity $\Delta h$ from the respective IDS-distributed 
    log file,
  \item compute the cartesian-to-topocentric rotation matrix $\bm{R}(\bm{r}\sp{\prime})$,
  \item add the topocentric eccentricity vector to get the antenna reference
    point coordinates for the beacon:
    \begin{equation}
      \bm{r} \equiv \bm{r}^{dpod}_{ARP} = 
        \bm{r}\sp{\prime} + \bm{R}^T \cdot 
          \begin{pmatrix} 0 & 0 & \Delta h \end{pmatrix}^T
      \label{eq:rg-to-rarp}
    \end{equation}
\end{enumerate}

\subsection{Coordinate \& Proper Time}
\label{ssec:coordinate-proper-time}
We use TAI as coordinate time; to transform RINEX observation time (given in 
proper time $\tau$) to coordinate time $t$, we use the \emph{receiver clock 
offset} values, extracted from the RINEX file (one value per observation block).

Hence, if an observation block is taged at proper time $\tau _i$ at the RINEX 
file, and the receiver clock offset for this block is $\Delta \tau _i$ (again from 
RINEX), then the coordinate time of the event in TAI is:
\begin{equation}
  t^{TAI}_i = \tau _i + \Delta \tau _i
\end{equation}

\subsection{Tropospheric Correction}
\label{ssec-tropospheric-correction}

\subsubsection{Mapping Functions}
We use the GPT3/VMF3 (\cite{Landskron2018}) to handle tropospheric refraction. 
Given the elevation angle $el$ and the beacon's ellipsoidal coordinates 
$\bm{r}=\begin{pmatrix} \lambda & \phi & h\end{pmatrix}$, we use the \texttt{gpt3} 
$\ang{5} \times \ang{5}$ grid to interpolate the $ah$ and $aw$ coefficients; 
we then compute the mapping function (with height correction) $mf_{wet}$ and 
$mf_{hydrostatic}$.

\subsubsection{Zenith Delays}
For the hydrostatic zenith delay $zd_{hydrostatic}$ we use the ``refined'' 
\emph{Saastamoinen} model (\cite{Davisetal85} and \cite{Saastamoinen72}). 
The corresponding value for the wet delay, $zd_{wet}$ is estimated during the 
analysis, \emph{per beacon and per pass}, using an initial value provided by 
\cite{Askneetal87}.

\subsubsection{Tropospheric Correction}
Using the mapping function and the zenith delay, we derive the tropospheric 
delay for an observation at $t=t_i$, given by:
\begin{equation}
  \delta _{TRO} [\si{\m}] = zd_{hydrostatic} \cdot mf_{hydrostatic} + zd_{wet}\at{t=t_i} \cdot mf_{wet}
  \label{eq:tropo-delay}
\end{equation}

The tropospheric correction term in \ref{eq:lem13}, is actually the ``time-differenced'' 
tropospheric delay between two measurements (the same ones used to derive the 
Doppler count), given in \si{\m \per \s}. That is:
\begin{equation}
  \begin{aligned}
    \Delta v_{TROPO} [\si{\m \per \s}] 
      &= \left( \delta _{TRO} \at{t=t_{i-1}} - \delta _{TRO} \at{t=t_{i}} \right) / \Delta \tau\\
      &= \left( \left[ zd_{h} \cdot mf_{h} + zd_{w}\at{t=t_{i-1}} \cdot mf_{w} \right]\at[\big]{t=t_i} - 
        \left[ zd_{h} \cdot mf_{h} + zd_{w}\at{t=t_{i-1}} \cdot mf_{w} \right]\at[\big]{t=t_{i-1}} \right) \ \Delta \tau
    \label{eq:dtrop-diff}
  \end{aligned}
\end{equation}

Note that in the above equation we use the same value $zd_{w}\at{t=t_{i-1}}$ 
for the wet part of the zenith delay, that is the best estimate prior to 
incorporating the (new) measurement at $t=t_i$.

The parameter $zd_{w}$ is estimated using no constraints and a simple white 
noise model (no process noise). Since it is an estimated parameter, we need 
to compute the (partial) derivative of the observation equation w.r.t to it, 
that is:
\begin{equation}
  \frac{\partial v_{theo}}{\partial zd_{w}} = \frac{mf_{w}\at[\big]{t=t_i} 
    - mf_{w}\at[\big]{t=t_{i-1}}}{\Delta \tau}
  \label{eq:partials-zwd}
\end{equation}

\subsection{Ionospheric Correction}
\label{ssec:iono-correction}
The basic observation equation \ref{eq:lem13}, is formed for the \SI{2}{\GHz} 
carrier. For each measurement, we have to account for the ionospheric path 
delay, by computing a correction (in cycles) as (\cite{lemoine-2016}, Sec. 2.5.7):
\begin{equation}
  \delta_{ION} [\SI{2}{\GHz}\text{ cycles}] = 
    \frac{L_{\SI{2}{\GHz}} - \sqrt{\gamma} \cdot L_{\SI{400}{\MHz}}}{\gamma - 1}
  \label{eq:iono-delay-cycles}
\end{equation}

which is added to the \SI{2}{\GHz} measurement at time $t=t_i$ (obtained by the 
RINEX file). Thus, the corrected observation is:
\begin{equation}
  L_{\SI{2}{\GHz},IF} [\SI{2}{\GHz}\text{ cycles}] = 
    L_{\SI{2}{\GHz}} + \delta_{ION}
  \label{eq:l2if}
\end{equation}

Note that after applying \ref{eq:l2if}, we should refer to the ``Iono-Free'' 
geometrical endpoints of the observations (and not the \SI{2}{\GHz} endpoints). 
This means that we have to apply respective \emph{phase center} corrections 
both at the satellite and at the beacon.

When applied to \ref{eq:lem13}, we have to ``differenciate'' the ionospheric 
delay computed from \ref{eq:iono-delay-cycles}, affecting two observations 
(the same ones used to derive the Doppler count). Hence, the term 
$\Delta v_{IONO}$ appearing in \ref{eq:lem13} is:
\begin{equation}
  \Delta v_{IONO} [\si{\m \per \s}] = 
    \tikz[baseline]{
      \node[fill=blue!20,anchor=base](en1){$\frac{c}{f_{e_N}}$};
    }
    \cdot 
    \frac{\delta_{ION}\at[\big]{t=t_{i-1}} 
    - \delta_{ION}\at[\big]{t=t_i}}{\Delta \tau}
  \label{eq:dion-diff}
\end{equation}

\begin{itemize}
  \item is this correct? should this be $f_{e_N}$?
    \tikz\node[fill=blue!20,draw,circle] (in1) {};
  \item maybe use the ``true'' emitter frequency, from 
  $f_{e_T} = f_{e_N} \cdot ( 1 + {\frac{\Delta f_e}{f_{e_N}}}\at[\big]{t=t_i})$ 
  but then we will have to re-arrange \ref{eq:lem13} so that the term 
  $\Delta v_{IONO}$ appears in the $v_{theo}$ part
\end{itemize}

\begin{tikzpicture}[overlay]
  \path[->] (in1) edge [bend left] (en1);
  %\path[->] (n2) edge [bend right] (t2);
  %\path[->] (n3) edge [out=0, in=-90] (t3);
\end{tikzpicture}

\subsubsection{2GHz \& Iono-Free Phase Center}
\label{sssec:2ghz-ionofree-pco}
To get to the iono-free phase center from the beacon \gls{arp}, 
we use (\cite{lemoine-2016}, Sec. 2.5.7, Eq. 20):
\begin{equation}
  \bm{r}_{iono-free} = \bm{r}_{\SI{2}{\GHz}} + \frac{\bm{r}_{\SI{2}{\GHz}} 
    - \bm{r}_{\SI{400}{\MHz}}}{\gamma - 1}
  \label{eq:ionf-pco}
\end{equation}

where $\bm{r}_{\SI{2}{\GHz}}$ and $\bm{r}_{\SI{400}{\MHz}}$ are the 
eccentricities of the \SI{2}{\GHz} and the \SI{400}{\MHz} carriers respectively 
from the beacon antena phase center (given eg at \cite{DORISGSM}, Sec. 5.2.1).

Note that in \ref{eq:ionf-pco}, the eccentricity vector $\bm{r}_{iono-free}$ 
is in a topocentric RF. Hence, to compute the ECEF coordinates of the 
iono-free phase center, given that we have the ECEF coordinates of the beacon's 
\gls{arp} $\bm{r}_{arp}$ (see \ref{ssec:beacon_coordinates}), we use:
\begin{equation}
  \bm{r}^{ecef}_{iono-free} = \bm{r}_{arp} + \bm{R}^T \cdot \bm{r}_{iono-free}
  \label{eq:arp-to-if-pc}
\end{equation}
where $\bm{R}$ is the cartesian-to-topocentric rotation matrix, computed 
at $\bm{r}_{arp}$.

In accordance to the beacons, a similar geometric reduction must be applied 
at the satellite's end, to correct for the discrepancy between the \SI{2}{\GHz} 
and the \emph{iono-free} phase center:
\begin{equation}
  \bm{r}^{satf}_{iono-free} = \bm{r}^{satf}_{\SI{2}{\GHz}} + 
    \frac{\bm{r}^{satf}_{\SI{2}{\GHz}} - 
    \bm{r}^{satf}_{\SI{400}{\MHz}}}{\gamma - 1}
\end{equation}

where the superscript $satf$ denotes the \emph{satellite-fixed} body/reference 
frame. On-board satellite antenna phase center offset values can be found in 
\cite{DorisSatModels}.

\subsubsection{Relativistic Correction}
\label{sssec:relativistic-correction}
\ref{eq:lem13} contains a relativistic correction term, $\Delta u_{REL}$. This 
correction is plit into two parts, $\Delta u_{REL_c}$ the part containing the 
clock correction and $\Delta u_{REL_r}$, containing the effect of the travel 
path. Here, we are only considering the $\Delta u_{REL_c}$ part.

The correction is computed (for a given obserbation) from the formula:
\begin{equation}
  \Delta u_{REL_c} = \frac{1}{c} \left( U_r - U_e + \frac{V^2_r - V^2_e}{2} \right) \si{\m}
  \label{eq:lem14}
\end{equation}
For the receiver part, the respective quantities in \ref{eq:lem14} are given by:
\begin{equation}
  \begin{aligned}
    V^2_r &= \norm{\bm{v}_{ecef}}^2 \\
    U_r   &= \frac{\mu}{\norm{\bm{r}_{ecef}}} \cdot \left( 1 - 
      \left(\frac{\alpha}{\norm{\bm{r}_{ecef}}}\right) ^2 \cdot J_2 \cdot
        \frac{3 \cdot \sin ^2{\phi} -1}{2} \right)
  \end{aligned}
  \label{eq:potential-receiver}
\end{equation}
where $\bm{r}_{ecef}$ and $\bm{v}_{ecef}$ are the position and velocity of the 
satellite at the given instant, in the terrestrial RF, aka ITRF.

For the emitter, things are further simplified. We consider $V_e = 0$ and the 
potential is computed as 
\begin{equation}
  U_e = \frac{\mu}{\norm{\bm{r}_{ecef}}}
  \label{eq:potential-emitter}
\end{equation}
where $\bm{r}_{ecef}$ is the position of the beacon in ITRF.

In \ref{eq:lem13}, we need the ``differentiated'' relativistic corrections, thus 
\begin{equation}
  \label{eq:drel-diff}
  \begin{aligned}
    \Delta v_{REL} &= \frac{1}{c} \left( \left[ U_r - U_e + \frac{V^2_r - V^2_e}{2} \right]\at{t=t_i} - \left[ U_r - U_e + \frac{V^2_r - V^2_e}{2} \right]\at{t=t_{i-1}} \right) \\
      &= \frac{1}{c} \cdot \left( U_r\at{t_i} - U_r\at{t_{i-1}} +  \frac{V_r\at{t_i}-V_r\at{t_{i-1}}}{2} \right) \si{\m\per\s}
  \end{aligned}
\end{equation}

\subsubsection{RINEX Observation Flags}
\label{sssec:rnx-flags}
An observation is rendered as ``non-usable'', if any of the below holds true:
\begin{itemize}
  \item flag1 or flag2 of the L\SI{2}{\giga\hertz} carrier is set to \texttt{1}
  \item flag1 or flag2 of the L\SI{400}{\mega\hertz} carrier is set to \texttt{1}
  \item flag1 of the Power (\texttt{W}) observation on either the L\SI{2}{\giga\hertz} 
    or the L\SI{400}{\mega\hertz} carrier is set to \texttt{1} (restart mode)
\end{itemize}

\subsubsection{Jason3 Quaternions}
\label{sssec:ja3-quaternions}
Currently we are only using the \emph{body-quaternion} files for Jason-3 
satellite. When we request a quaternion for a given epoch $t$, then we interpolate 
using the \texttt{slerp} method. 

We assume that the recorded quaternions, act in the sense:
\begin{equation}
  \bm{x}_{sv-fixed} = q(t_i) \otimes \bm{x}_{gcrs}
\end{equation}

The most often operation used involving the quaternions, is transforming the 
satellite position vector from satellite's center of gravity (CoG) to the 
DORIS antenna reference point (ARP). If the satellite CoG vector in the GCRS 
is $\bm{x}^{cog}_{gcrs}$, then the vector to the ARP is:
\begin{equation}
  \bm{x}^{arp}_{gcrs} = \tilde{q}(t_i) \otimes (\bm{x}^{arp}_{sv} - \bm{x}^{cog}_{sv})
\end{equation}
where $\tilde{q}$ is the conjugate of $q$ and the subscript $sv$ denotes vectors 
wrt the satellite-fixed reference frame.

\subsection{Filtering}
\label{ssec:filtering}
The estimated parameter set used in the Kalman filter is:
\begin{equation}
  \bm{X} = \begin{pmatrix}
    \begin{pmatrix} r_x & r_y & r_z & v_x & v_y & v_z \end{pmatrix}^T_{ecef} \\
    C_{drag} \\
    \begin{pmatrix} L^z_w & F_0 & F_1 \end{pmatrix}^T_{beacon_1} \\
    \vdots \\
    \begin{pmatrix} L^z_w & F_0 & F_1 \end{pmatrix}^T_{beacon_N} \\
  \end{pmatrix}
\end{equation}
where $F_0$ and $F_1$ are the coefficients for modeling the bias and drift of the 
relative frequency offset of (each) emitter.

We need the partials of \ref{eq:lem13b} wrt $X$:
\begin{equation}
  \frac{\partial v_{theo}}{\partial \bm{x}_{ecef}} = \bm{R}^T \left( \frac{\bm{s}}{\norm{\bm{s}}}\at{t=t_i} - \frac{\bm{s}}{\norm{\bm{s}}}\at{t=t_{i-1}} \right) / \Delta \tau
  \label{eq:partials-pos}
\end{equation}
where $\bm{x}_{ecef}$ is the ECEF position of the satellite, $\bm{R}$ is the 
cartesian-to-topocentric matrix (centered at the beacon site), $\bm{s}$ is the 
topocentric beacon-to-satellite vector.

For the satellite velocity $\bm{v}_{ecef}$, we have
\begin{equation}
  \frac{\partial v_{theo}}{\partial \bm{v}_{ecef}} = \bm{0}
  \label{eq:partials-vel}
\end{equation}

\section{Processing SetUp}

\subsubsection{General SetUp}
\begin{itemize}
  \item Elevation cut-off angle \ang{10}
  \item Setup new satellite arc if previous observation (of the beacon) is more 
    than \SI{5}{\min} away
  \item First satellite state (for reference time of RINEX) extracted from a relevant 
    sp3 file
  \item EOP information is extracted/interpolated from IERS \texttt{C04} files.
  \item Using gravity model \texttt{GOCO02s} (icgem format), up to degre/order 90.
  \item To compute atmospheric density, we are using the \texttt{NRLMSISE00} model, 
    using flux data from CelesTrack.
  \item Using JPL's \texttt{de405} ephemeris to compute Sun and Moon positions.
  \item Satellite information \texttt{ja3mass.txt} and \texttt{DORISSatelliteModels.pdf} 
\end{itemize}

\subsubsection{Filter SetUp}
\begin{itemize}
  \item A-priori information for $\Delta f_e / f_{e_N}$: $0.0 \pm 1.0$
  \item A-priori information for drag coefficient $C_{d}$: $2.0 \pm .75$
  \item A-priori value for $L^z_w$ computed from Aske \& Nordius (recommended by VMF3) $\pm 0.5 \si{\m}$
\end{itemize}

%\subsubsection{Beacon Information}
%Reference coordinates extracted from \texttt{dpod2014\_053.snx}. Antenna 
%Reference Point (eccentricity) extracted from respective IDS log files, and 
%added to the extrapolated cartesian coordinates (see 
%\ref{ssec:beacon_coordinates}).

%\subsubsection{Satellite Information}
%Centre of Gravity (CoG) coordinates, in the SV-fixed RF, are extracted from the 
%file \texttt{ja3mass.txt}. Phase centre coordinates (in the SV-fixed RF) for the 
%two frequencies are extracted from the file \texttt{DORISSatelliteModels.pdf} and 
%converted to \emph{iono-free} pco as:
%\begin{equation}
%  \bm{r}_{iono-free-pco} = \bm{r}_{2GHz,PCO} + (\bm{r}_{400MHz,PCO} - \bm{r}_{2GHz,PCO}) / (\gamma - 1)
%\end{equation}
%with all $\bm{r}$ vectors wrt the the SV-fixed RF.

\subsubsection{Algorithm}
\begin{enumerate}
  \item Setup auxiliary infromation, data files, filter, etc ...
  \begin{itemize}
    \item Extrapolate coordinates for every beacon to RINEX reference epoch using \ref{eq:rg-to-rarp}, see \ref{ssec:beacon_coordinates}
  \end{itemize}

  \item Parse the RINEX file header

  \item \textbf{For every observation block in the RINEX file}:
    \begin{enumerate}\label{every-block}
      \item Transform proper time $\tau _i$ to TAI $t_i$ using the receiver-clock-offset 
        value reported in the RINEX
      \item Integrate orbit to $t_i$ (results are state vector $\bm{X}_{t_i}$ and state 
        transition matrix $\bm{\Phi}_{t_i}$ in GCRS)
      \item Perform the Kalman filter time-update, aka compute new $P$ matrix as 
        \begin{equation}
          P^- _{t_i} = \bm{\Phi} \cdot P^+ _{t_{i-1}} \cdot \bm{\Phi}^T
        \end{equation}
        where $\bm{\Phi} = \begin{pmatrix} \bm{\Phi}_{6\times 6} & \bm{0} \\ \bm{0} & \bm{I} \end{pmatrix}$
        \item Compute SV-coordinates of Iono-Free Phse Center, in ITRF aka $\bm{X}^{ecef}_{iono-free,t_i}$
        \item \textbf{For every observation set in the current block}:
          \begin{enumerate}\label{every-observation}
            \item Check observation flags reported in the RINEX file (see \ref{sssec:rnx-flags}). 
              If the observation is un-usable, then mark a discontinuity (for the beacon) and procceed 
              to next observation (goto \ref{every-observation}). If flags are ok, continue.
            \item Compute the Iono-Free Phase Center of the beacon using \ref{eq:arp-to-if-pc} 
              (see \ref{sssec:2ghz-ionofree-pco}) to get $\bm{r}^{ecef}_{iono-free}$
            \item \label{cmp-rho} Transform the satellite-to-beacon vector from ITRF to topocentric coordinates, 
              and compute azimouth $Az$, elevation $El$ and range $s$ (use $\bm{X}^{ecef}_{iono-free,t_i}$ and 
              $\bm{r}^{ecef}_{iono-free}$ so that the vector is wrt the Iono-Free 
              Phase center at both ends).
            \item If $El < \text{ minimum Elevation Angle}$ goto \ref{every-observation}; else procceed
            \item Check if this is the first observation of the beacon, or we have to start 
              a new arc, i.e. previous observation of the beacons happened more than 
              \SI{5}{\min} away
            \item Compute beacon's nominal frequency $f_{e_N}$, via \ref{eq:nominal-freq} 
            \item Compute ionospheric correction $\Delta I_{t_i}$ (in cycles) using \ref{eq:iono-delay-cycles}
            \item Compute tropospheric delay for the observation $\Delta T_{t_i}$, 
              using \ref{eq:tropo-delay}. The value $zd_{wet}$ is the last estimated 
              value, aka $zd_{wet}\at{t=t_{i-1}}$
            \item Compute relativistic correction using $\Delta R_{t_i}$ using \ref{eq:lem14}
              (use \ref{eq:potential-receiver} and \ref{eq:potential-emitter}; values 
              $\mu$, $\alpha$ and $J_2$ needed to compute \ref{eq:potential-receiver} 
              are extracted from the gravity model
            \item Check for discontinuity or start of new arc; if true, then store 
              observation and goto \ref{every-observation}. Else retrieve last 
              stored dataset (latest beacon observation)
            \item Compute the true proper frequency of the receiver, $f_{r_T}\at{t=t_i}$ 
              using \ref{eq:tpf-rec} (scaled with $10^6$ to get \si{\Hz})
            \item Compute the Doppler count $NDpod$, by just subtracting the current 
              \SI{2}{\GHz} carrier observation with the previous one, aka
              $NDop = L_{\SI{2}{\GHz}}\at{t=t_i} - L_{\SI{2}{\GHz}}\at{t=t_{i-1}}$
              as recorded in the RINEX file
            \item Differentiate corrections wrt to (\textbf{proper}) time:
              \begin{itemize}
                \item $\Delta v_{ION}$ from \ref{eq:dion-diff}
                \item $\Delta v_{TRO}$ from \ref{eq:dtrop-diff}
                \item $\Delta v_{REC}$ from \ref{eq:drel-diff}
              \end{itemize}
            \item Compute observed value $v_{measured}$ (range-rate) via \ref{eq:lem13a}
            \item Compute ``model'' value $v_{theo}$ via \ref{eq:lem13b}. Note that the ranges $\rho _2$ and 
              $\rho _1$ are the norms of the topocentric beacon-to-satellite vectors computed in 
              step \ref{cmp-rho} at current and previous measurement, aka $\rho _2 \equiv \bm{s}_{t_i}$ 
              and $\rho _2 \equiv \bm{s}_{t_{i-1}}$. The term $\frac{\Delta f_e}{f_{e_N}}$ is the 
              best current estimate (from the filter)
            \item Compute the residual value $u = v_{measured} - v_{theo}$ and 
              perform a ``harsh'' $3\sigma$ test. If $ |u| > 3 \sigma / \sin{El}$ 
              goto \ref{every-observation}, else continue
            \item Compute partials wrt the estimated parameters:
              \begin{itemize}
                \item partials wrt satellite position vector: 
                \begin{equation} 
                  \frac{\partial v_{theo}}{\partial \bm{X}^{ecef}_{iono-free,t_i}} = 
                    \bm{R}^T \cdot \frac{1}{\Delta \tau} \begin{pmatrix} 
                      \frac{\bm{s}_{t_i}}{\norm{\bm{s}_{t_i}}} - \frac{\bm{s}_{t_{i-1}}}{\norm{\bm{s}_{t_{i-1}}}}
                      \end{pmatrix}
                \end{equation}
                where $\bm{R}$ is the cartesian-to-topocentric matrix and $\bm{s}$ being the 
                beacon-to-satellite topocentric vector
                \item partials wrt satellite velocity: 
                \begin{equation} 
                  \frac{\partial v_{theo}}{\partial \bm{v}^{ecef}_{iono-free,t_i}} = \bm{0}
                \end{equation}
                \item partials wrt drag coefficient:
                \begin{equation}
                  \frac{\partial v_{theo}}{\partial C_{drag}} = 1
                \end{equation}
                \item partials wrt emmiter relative frequency offset: note that 
                $\Delta f_e / f_{e_N}$ is modeled as a linear model, hence
                $\Delta f_e / f_{e_N}\at{t_i} \equiv \Delta F(t_i) = F_1 + F_2 \cdot (t_i - t_0)$, so that
                \begin{equation}
                  \begin{aligned}
                    \frac{\partial v_{theo}}{\partial F_1} &= 
                      -\frac{c(\frac{N_{DOP}}{\Delta\tau_r} + f_{r_T})}{f_{e_N}}\\
                    \frac{\partial v_{theo}}{\partial F_2} &= 
                      -\delta \tau \cdot \frac{c(\frac{N_{DOP}}{\Delta\tau_r} + f_{r_T})}{f_{e_N}} 
                  \end{aligned}
                \end{equation}
                \item partials wrt zenith wet delay from \ref{eq:partials-zwd}
              \end{itemize}
            \item Perform filter measurement update, weighting the observation with 
              $\sigma _{obs} = \sigma_0 / \cos{El}$. That will result in a new 
              ``corrected'' satellite state vector $\bm{X}^{ecef}_{iono-free}$; repeat 
              loop, using this updated state, for the rest of the beacons (aka goto \ref{every-observation})
          \end{enumerate}
      \end{enumerate}
\end{enumerate}
